---
title: "IPS9 in R: Nonparametric tests (Chapter 15)"
author: "Shukry Zablah (szablah20@amherst.edu) and Nicholas Horton (nhorton@amherst.edu)"
date: "July 21, 2018"
output: 
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 4
    fig_width: 6
---


```{r, include = FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include = FALSE}
# knitr settings to control how R chunks work.
knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed
  size = "small"    # slightly smaller font for code
)
```

## Introduction and background 

These documents are intended to help describe how to undertake analyses introduced 
as examples in the Ninth Edition of \emph{Introduction to the Practice of Statistics} (2017) by Moore, McCabe, and Craig.

More information about the book can be found [here](https://macmillanlearning.com/Catalog/product/introductiontothepracticeofstatistics-ninthedition-moore).
The data used in these documents can be found under Data Sets in the [Student Site](https://www.macmillanlearning.com/catalog/studentresources/ips9e?_ga=2.29224888.526668012.1531487989-1209447309.1529940008#). This
file as well as the associated R Markdown reproducible analysis source file used to create it can be found at https://nhorton.people.amherst.edu/ips9/.

This work leverages initiatives undertaken by Project MOSAIC (http://www.mosaic-web.org), an NSF-funded effort to improve the teaching of statistics, calculus, science and computing in the undergraduate curriculum. In particular, we utilize the `mosaic` package, which was written to simplify the use of R for introductory statistics courses. A short summary of the R needed to teach introductory statistics can be found in the mosaic package vignettes (http://cran.r-project.org/web/packages/mosaic).  A paper describing the mosaic approach was published in the *R Journal*: https://journal.r-project.org/archive/2017/RJ-2017-024.
  
## Chapter 15: Nonparametric tests
This file replicates the analyses from Chapter 15: Nonparametric tests.

First, load the packages that will be needed for this document: 
```{r load-packages, message=FALSE}
library(mosaic)
library(readr)
library(tidyr)
```

The skills in the this chapter will help analyze data that doesn't follow the Normal distribution and is not fit for the tests we previously looked at. 

### Section 15.1: The Wilcoxon rank sum test

Let's read in the csv file for the Hits data in Example 15.1. 

```{r Hits1, message=FALSE}
Hits <- read_csv("https://nhorton.people.amherst.edu/ips9/data/chapter15/EG15-001HITS.csv")
Hits
```

To recreate the table shown, just select the two relevant columns for the hits of each group and drop the `NA` values. 

```{r Hits2}
Hits %>% 
  select(HitsAmer, HitsNat) %>%
  tidyr::drop_na()
```

Dealing with `NA` values inside our observations can take up uneccesary space. Use the `tidyr::drop_na()` function to remove observations containing any `NA` values from your dataset.  

Now let's set up the rank transformation. 

```{r Hits3}
nrows <- nrow(Hits)
RanksHits <- Hits %>%
  select(League, Hits) %>%
  arrange(Hits) %>%
  mutate(Rank = 1:nrows)
RanksHits
```

Next we sum the ranks for each group. We will take advantage of the `group_by()` and `summary()` idiom.

```{r Hits4}
RanksHits %>%
  group_by(League) %>%
  summarize(`Sum of ranks` = sum(Rank))
```

The intuition behind the test is that if there was no difference between the leagues both sums would be the same. 

Finally, to perform the Wilcoxon Rank Sums test we use `wilcox.test()`.

```{r Hits5}
wilcox.test(Hits ~ League, data = Hits)
```

In Example 15.6 the book discusses how to deal with ties in the dataset (remember we rank the observations).

We have to recreate the dataset and confirm we can get the table of counts. 

```{r Exerg1}
Exerg <- rbind(
  do(6) * data.frame(Exergamer = "Yes", TV_time = "None"),
  do(160) * data.frame(Exergamer= "Yes", TV_time = "<2 hours"),
  do(115) * data.frame(Exergamer = "Yes", TV_time = ">=2 hours"),
  do(48) * data.frame(Exergamer = "No", TV_time = "None"),
  do(616) * data.frame(Exergamer = "No", TV_time = "<2 hours"),
  do(255) * data.frame(Exergamer = "No", TV_time = ">=2 hours")
)
```

```{r Exerg2}
tally(~ Exergamer + TV_time, data = Exerg)
tally(~ Exergamer + TV_time, data = Exerg, format = "percent", margins = TRUE)
```

Sucess! Now let's perform the Wilcox Rank Sum Test. R will take care of the fact that there are "ties" in the dataset. However, we do have to code the variables into numerics in order for the test to run. 

```{r Exerg3}
Exerg %>% 
  mutate(Val = recode(TV_time,
                      `None` = 0,
                      `<2 hours` = 1,
                      `>=2 hours` = 2)
         ) %>% 
  wilcox.test(Val ~ Exergamer, data = .)
```


### Section 15.2: The Wilcoxon signed rank test

Read in the csv file. 

```{r Story1, message=FALSE}
Story <- read_csv("https://nhorton.people.amherst.edu/ips9/data/chapter15/EG15-009STORY.csv")
Story
```

Filter the relevant observations by using `filter()`.

```{r}
StoryTable <- Story %>%
  filter(Progress == "Low") 
```

We will prove why our previous tests are not suitable for the scenario. 

```{r}
gf_qq(~ Difference, data = StoryTable)
gf_histogram(~ Difference, data = StoryTable, binwidth = 0.2, center = 0.5)
```

The plots suggest lack of Normality. This is why we use a rank test. 

In R, the same `wilcox.test()` function will be used. However, we specify `paired = TRUE` in order to perform the signed rank test from the book. The option `alternative = "less"` gets the one sided p-value that we are testing for. 

```{r}
wilcox.test(Values ~ Stories, data = x, paired = TRUE, alternative = "less")
```

Example 15.12 uses the same function as before. 

### Section 15.3: The Kruskal-Wallis test
