---
title: "IPS9 in R: Looking at Data -- Relations (Chapter 2)"
author: "Margaret Chien and Nicholas Horton (nhorton@amherst.edu)"
date: "July 13, 2018"
output: 
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 4
    fig_width: 6
---


```{r, include = FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include = FALSE}
# knitr settings to control how R chunks work.
knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed
  size = "small"    # slightly smaller font for code
)
```

## Introduction and background 

These documents are intended to help describe how to undertake analyses introduced 
as examples in the Ninth Edition of \emph{Introduction to the Practice of Statistics} (2017) by Moore, McCabe, and Craig.

More information about the book can be found [here](https://macmillanlearning.com/Catalog/product/introductiontothepracticeofstatistics-ninthedition-moore).
The data used in these documents can be found under Data Sets in the [Student Site](https://www.macmillanlearning.com/catalog/studentresources/ips9e?_ga=2.29224888.526668012.1531487989-1209447309.1529940008#). This
file as well as the associated R Markdown reproducible analysis source file used to create it can be found at https://nhorton.people.amherst.edu/ips9/.

This work leverages initiatives undertaken by Project MOSAIC (http://www.mosaic-web.org), an NSF-funded effort to improve the teaching of statistics, calculus, science and computing in the undergraduate curriculum. In particular, we utilize the `mosaic` package, which was written to simplify the use of R for introductory statistics courses. A short summary of the R needed to teach introductory statistics can be found in the mosaic package vignettes (http://cran.r-project.org/web/packages/mosaic).  A paper describing the mosaic approach was published in the *R Journal*: https://journal.r-project.org/archive/2017/RJ-2017-024.
  
## Chapter 2: Looking at Data -- Relationships
This file replicates the analyses from 2: Looking at Data -- Relationships.

First, load the packages that will be needed for this document: 
```{r load-packages}
library(mosaic)
library(readr)
```

### Section 2.1: Relationships 

### Section 2.2: Scatterplots

#### Example 2.8: Laundry detergents

```{r}
Laundry <- read_csv("https://nhorton.people.amherst.edu/ips9/data/chapter02/EG02-008LAUNDRY.csv")
```

```{r}
# 2.10: Examine the spreadsheet
str(Laundry)
favstats(~ Rating, data = Laundry)
gf_histogram(~ Rating, data = Laundry, binwidth = 5)
favstats(~ Price, data = Laundry)
gf_histogram(~ Price, data = Laundry, binwidth = 5)
```

#### Example 2.9: Laundry detergents

```{r}
Laundry <- read_csv("https://nhorton.people.amherst.edu/ips9/data/chapter02/EG02-009LAUNDRY.csv")
# Figure 2.1, page 86
gf_point(Rating ~ Price, data = Laundry, xlab = "Price per load (cents)")
# Figure 2.2
gf_point(Rating ~ Price, data = Laundry, xlab = "Price per load (cents)") +
  xlim(0, 35)
```

```{r}
# 2.12: Make a scatterplot
gf_point(jitter(Rating) ~ jitter(Price), data = Laundry, xlab = "Price per load (cents)") +
  xlim(0, 35)
```

We can use `jitter()` to add some noise into the plot to show overlapped points.

```{r}
# 2.13: Change the units
Laundry2 <- Laundry %>%
  mutate(Price = Price/100)
favstats(~ Price, data = Laundry2)
gf_point(Rating ~ Price, data = Laundry2, xlab = "Price per load (dollars)")
```

#### Example 2.10: Scatterplot with a straight line

```{r}
Laundry <- read_csv("https://nhorton.people.amherst.edu/ips9/data/chapter02/EG02-010LAUND.csv")
gf_point(Rating ~ Price, data = Laundry, xlab = "Price per load (cents)") %>%
  gf_lm() 
```

#### Example 2.11: Education spending and population: Benchmarking

```{r}
EduSpending <- read_csv("https://nhorton.people.amherst.edu/ips9/data/chapter02/EG02-011EDSPEND.csv")
head(EduSpending)
# Figure 2.5, page 90
gf_point(Spending ~ Population, data = EduSpending) %>%
  gf_lm() %>%
  gf_labs(x = "Population (millions)", y = "Spending ($ billion)")
```

#### Example 2.12: Calcium retention

```{r}
Calcium <- read_csv("https://nhorton.people.amherst.edu/ips9/data/chapter02/EG02-012CALCIUM.csv")
# Figure 2.6
gf_point(CaRetention~ CaIntake, data = Calcium) %>%
  gf_smooth() %>%
  gf_labs(x = "Calcium intake (mg/d)", y = "Calcium retention (mg/d)")
```

#### Example 2.13: Calcium retention with logarithms

```{r}
Calcium <- read_csv("https://nhorton.people.amherst.edu/ips9/data/chapter02/EG02-013CALCIUM.csv")
# Figure 2.7, page 91
gf_point(LogRet ~ CaIntake, data = Calcium) %>%
  gf_lm() %>%
  gf_labs(x = "Calcium intake (mg/d)", y = "Log calcium retention")
```

#### Example 2.14: Education spending and population with logarithms

```{r}
EduSpending <- read_csv("https://nhorton.people.amherst.edu/ips9/data/chapter02/EG02-014EDSPEND.csv")
# Figure 2.8, page 92
EduSpending %>%
  mutate(LogPop = log(Population), LogSpend = log(Spending)) %>%
  gf_point(LogSpend ~ LogPop) %>%
  gf_lm() %>%
  gf_labs(x = "Log population", y = "Log spending")
```

#### Example 2.15: Rating versus price and type of laundry detergent

```{r}
Laundry <- read_csv("https://nhorton.people.amherst.edu/ips9/data/chapter02/EG02-015LAUND.csv")
# Figure 2.9, page 93
gf_point(Rating ~ Price, color = ~ Type, data = Laundry, xlab = "Price per load (cents)") %>%
  gf_lm()
```

#### Example 2.16: Laundry rating versus price with a smooth fit

### Section 2.3: Correlation
### Section 2.4: Least Squares Regression
### Section 2.5: Cautions about Correlation and Regression
### Section 2.7: Data Analysis fro Two Way Tables
### Section 2.6: The Questionn of Causation


