---
title: "IPS9 in R: Probability: The Study of Randomness (Chapter 4)"
author: "Bonnie Lin and Nicholas Horton (nhorton@amherst.edu)"
date: "July 13, 2018"
output: 
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 4
    fig_width: 6
---


```{r, include = FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include = FALSE}
# knitr settings to control how R chunks work.
knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed
  size = "small"    # slightly smaller font for code
)
```

## Introduction and background 

These documents are intended to help describe how to undertake analyses introduced 
as examples in the Ninth Edition of \emph{Introduction to the Practice of Statistics} (2017) by Moore, McCabe, and Craig.

More information about the book can be found at 
The data used in these documents can be found under Data Sets in the [Student Site](https://www.macmillanlearning.com/catalog/studentresources/ips9e?_ga=2.29224888.526668012.1531487989-1209447309.1529940008#). This
file as well as the associated R Markdown reproducible analysis source file used to create it can be found at https://nhorton.people.amherst.edu/ips9/.

This work leverages initiatives undertaken by Project MOSAIC (http://www.mosaic-web.org), an NSF-funded effort to improve the teaching of statistics, calculus, science and computing in the undergraduate curriculum. In particular, we utilize the `mosaic` package, which was written to simplify the use of R for introductory statistics courses. A short summary of the R needed to teach introductory statistics can be found in the mosaic package vignettes (http://cran.r-project.org/web/packages/mosaic).  A paper describing the mosaic approach was published in the *R Journal*: https://journal.r-project.org/archive/2017/RJ-2017-024.
  
## Chapter 4: Probability: The Study of Randomness
This file replicates the analyses from Chapter 4: Probability: The Study of Randomness.

First, load the package that will be needed for this document: 
```{r load-package}
library(mosaic)
```

### Section 4.1: Randomness
The results of two trials of a coin toss simulation 5000 times
are plotted as the proportion of heads, as shown in Figure 3.1 (page 216). 
We can recreate the plot by typing: 
```{r eg4-1}
## XX should I set seed? and how do I plot the realization of two seeds on same graph? 
## XX can create a new column, and/or create a new dataset
tosses <- rbinom(1:5000, 1, 0.5)
x <- seq(1:5000)
cy <- cumsum(tosses)
phead <- (cy / x)
gf_line(phead ~ x, color = "red") %>% 
  gf_labs(x = "Number of tosses", y = "Proportion of heads") %>% 
  gf_hline(., yintercept = 0.5)
```
XX Should I specify the arguments in this function? 

The `rbinom()` function will ... 

The R function, `runif()`, generates a random number between 0 and 1. 
We can demonstrate the code below: 
```{r eg4-7}
runif(1)
```
Since the default arguments in the function define the sample space to be all numbers 
between 0 and 1, all we need to specify the number of random numbers we want outputted.
Run the code above several times. Notice that every iteration gives you a different output. 
If you do not set a seed, every time you run the code, you will get a random number. 
```{r}
runif(2) #change the number of random numbers generated from this sample space

set.seed(2018) #setting the seed to get reproducible results
runif(1)
```
Now that we have selected a pseudorandom seed, anyone who runs this code should expect to see
the output of `runif(1)` to be 0.34. 

### Section 4.2: 
