---
title: "IPS9 in R: Bootstrap Methods and Permutation Tests (Chapter 16)"
author: "Bonnie Lin and Nicholas Horton (nhorton@amherst.edu)"
date: "June 28, 2018"
output: 
  pdf_document:
    fig_height: 4
    fig_width: 6
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 4
    fig_width: 6
---


```{r, include = FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
library(readr)
library(janitor)
```

```{r, include = FALSE, message = FALSE}
# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy = FALSE,     # display code as typed
  size = "small",    # slightly smaller font for code
  message = FALSE
)
```

## Introduction and background 

This document is intended to help describe how to undertake analyses introduced 
as examples in the Ninth Edition of \emph{Introduction to the Practice of Statistics} (2017) by Moore, McCabe, and Craig.

More information about the book can be found at https://macmillanlearning.com/Catalog/product/introductiontothepracticeofstatistics-ninthedition-moore. This
file as well as the associated R Markdown reproducible analysis source file used to create it can be found at https://nhorton.people.amherst.edu/ips9/.

This work leverages initiatives undertaken by Project MOSAIC (http://www.mosaic-web.org), an NSF-funded effort to improve the teaching of statistics, calculus, science and computing in the undergraduate curriculum. In particular, we utilize the `mosaic` package, which was written to simplify the use of R for introductory statistics courses. A short summary of the R needed to teach introductory statistics can be found in the mosaic package vignettes (http://cran.r-project.org/web/packages/mosaic).  A paper describing the mosaic approach was published in the *R Journal*: https://journal.r-project.org/archive/2017/RJ-2017-024.
  
## Chapter 16: Bootstrap methods and permutation tests
The specific goal of this document is to demonstrate how to replicate the
analysis described in Chapter 16: Bootstrap methods and permutation tests, one of 
the four online companion chapters. 

### Section 16.1: The bootstrap idea
```{r load-packages}
library(mosaic)
library(readr)
library(janitor)
```
With the following code, we create many resamples by repeatedly 
sampling with replacement from the sample of 21 profiles. Then, we 
create the bootstrap distribution by finding the mean for the 3,000 
resamples. 
```{r eg 16-2}
FACE4 <- read_csv("https://nhorton.people.amherst.edu/ips9/data/chapter16/EG16-01FACE4.csv")

FACE4 <- FACE4 %>%
  mutate(Time = as.numeric(Time))
## XX class() denotes Time as numeric but the message suggests that column is specified as double

set.seed(2018)
numsim <- 3000
Avg_Viewtime <- do(numsim) * mean(~ Time, data = resample(FACE4, 21)) 

#Figure 16.3(a)
mean_time = FACE4 %>%
  pull(Time) %>%
  mean()

gf_histogram(~ mean, data = Avg_Viewtime) %>%
  gf_labs(x = "Mean time of resamples (in minutes)", y = "Count") %>%
  gf_vline(x = mean_time) 

#Figure 16.3(b)
gf_qq(~ mean, data = Avg_Viewtime) %>% 
  gf_qqline(color = "red", linetype = "solid") %>% 
  gf_labs(x = "Normal score", y = "Mean time of resamples (in minutes)")
```
The psuedorandom number used to set the seed enables the reproduction of results. 
The `do()` function in the `mosaic` package simplifies the iteration and resampling  
processees by allowing us to set the number of repetitions and specify what to repeat. 

### Section 16.2: First steps in using the bootstrap
```{r eg 16-4}
College_GPA <- read_csv("https://nhorton.people.amherst.edu/ips9/data/chapter16/EG16-04GPA.csv")

## Figure 16.8(a)
gf_dhistogram(~ GPA, data = College_GPA) %>%
  gf_labs(x = "GPA", y = "Percent") 

## Figure 16.8(b)
gf_qq(~ mean, data = Avg_Viewtime) %>% 
  gf_qqline(color = "red", linetype = "solid") %>% 
  gf_labs(x = "Normal score", y = "Mean time of resamples (in minutes)")
```
### Section 16.3: How accurate is a bootstrap distribution?
### Section 16.4: Bootstrap confidence intervals 
### Section 16.5: Significance testing using permutation tests

