---
title: "IPS9 in R: One-way analysis of variance (Chapter 12)"
author: "Shukry Zablah (szablah20@amherst.edu) and Nicholas Horton (nhorton@amherst.edu)"
date: "July 19, 2018"
output: 
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 4
    fig_width: 6
---


```{r, include = FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include = FALSE}
# knitr settings to control how R chunks work.
knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed
  size = "small"    # slightly smaller font for code
)
```

## Introduction and background 

These documents are intended to help describe how to undertake analyses introduced 
as examples in the Ninth Edition of \emph{Introduction to the Practice of Statistics} (2017) by Moore, McCabe, and Craig.

More information about the book can be found [here](https://macmillanlearning.com/Catalog/product/introductiontothepracticeofstatistics-ninthedition-moore).
The data used in these documents can be found under Data Sets in the [Student Site](https://www.macmillanlearning.com/catalog/studentresources/ips9e?_ga=2.29224888.526668012.1531487989-1209447309.1529940008#). This
file as well as the associated R Markdown reproducible analysis source file used to create it can be found at https://nhorton.people.amherst.edu/ips9/.

This work leverages initiatives undertaken by Project MOSAIC (http://www.mosaic-web.org), an NSF-funded effort to improve the teaching of statistics, calculus, science and computing in the undergraduate curriculum. In particular, we utilize the `mosaic` package, which was written to simplify the use of R for introductory statistics courses. A short summary of the R needed to teach introductory statistics can be found in the mosaic package vignettes (http://cran.r-project.org/web/packages/mosaic).  A paper describing the mosaic approach was published in the *R Journal*: https://journal.r-project.org/archive/2017/RJ-2017-024.
  
## Chapter 12: One-way analysis of variance
This file replicates the analyses from Chapter 12: One-way analysis of variance.

First, load the packages that will be needed for this document: 
```{r load-packages}
library(mosaic)
library(readr)
```

> Complicated computations do not guarantee a valid statistical analysis. (648)

### Section 12.1: Inference for one-way analysis of variance

We begin with Example 12.3 in page 648. Let's read in our data

```{r Friends1, message=FALSE}
Friends <- read_csv("https://nhorton.people.amherst.edu/ips9/data/chapter12/EG12-03FRIENDS.csv")
Friends <- Friends %>%
  mutate(Friends = as.factor(Friends))
head(Friends)
```

We want to get a nice summary table like the one in example 12.3. To do this we have to group the dataset by the number of friends. Then we can use some aggregating functions to get the desired table structure. 

```{r Friends2}
SummaryFriends <- Friends %>%
  group_by(Friends) %>%
  summarize(n = n(), x_hat = mean(Score), s = sd(Score))
SummaryFriends
```

In R the combination of `group_by()` and `summary()` is a powerful one. With these two data verbs we can create summary outputs like the one above with very little code.

There are some helpful visualizations starting in page 649 that will help us in our analysis of our Friends data. 

To recreate the histogram (note the `|` used to facet based on the number of friends): 

```{r Friends3}
gf_histogram(~ Score | Friends, data = Friends) #XX need to make this match the figure
```

To recreate the boxplot: 

```{r Friends4}
gf_boxplot(Score ~ Friends, data = Friends)
```

And to recreate the linegraph (note the piping through gf_point to add the points to the line): 

```{r Friends5}
gf_line(x_hat ~ as.numeric(Friends), data = SummaryFriends) %>% gf_point() 
```

These visualizations are a quick way to know what is going on with your data. Just a few lines of codes can save much time in your analysis. 

```{r}
anova(lm(Score ~ Friends, data = Friends))
```

### Section 12.2: Comparing the means
